---
title: "Introduction"
author: "Lian Rui"
date: "2024-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR2)
```


# An Overview of Statistical Learning

## Wage Data

Regression, linear/non-linear multivariate, variation unexplained. 

## Stock Market Data

```{r}
data(Smarket)

# Fit a logistic regression model using lagged returns
glm.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data=Smarket, family=binomial)

# Summarize the model
summary(glm.fit)

```


## Gene Expression Data

PCA

# A Brief History of Statistical Learning

# This Book

# Who Should Read This Book

# Organization of This Book

Ch2: 

  - Basic terminology and concept
  
  - KNN
  
Ch3--Ch4: 

  - Linear method: regression and classification
  
  - Logistic regression and LDA
  
Ch5: 

  - Cross validation and bootstrap
  
Ch6: 

 - Stepwise selection
 
 - Ridge regression
 
 - Principal component regression
 
 - Lasso
 
Ch7: 

  - Additive model
  
Ch8: 

  - Bagging
  
  - Boosting
  
  - Random Forest
  
  - Support vector machines
  
Ch9: 

  - Deep learning
  
Ch10-11: 

  - Survival analysis
  
Ch12: 

 - PCA
 
 - Kmeans
 
 - Hierarchical clusting



